Can Symmetric Cryptography Be Liberated from the Von Neumann Style?

### 2024-10-12 paper.tex
Rephrase

We have thus completed our journey toward a rationalized treatment of
bitslicing and fixslicing in their quintessential form, taking the
\Skinny{} cipher as our running example.
%
%% * Other applicative: Gift
%% ** Specification: only thinking: Rows.T Cols.T Slice4.T bool
%% *** rest is rote functionalisation
%% ** Bitslicing: extract Slice4.T out and Double
%% ** Fixslicing: decompose permutation into in-register followed by 90 degree rotation
%% *** cancel out after 4
%% *** -> simpler
%
We have similarly treated the \Gift{} cipher, whose design is in fact at
the origin of the fixslicing
technique~\citep{adominicai:fixslicing-gift}. \Gift{} is specified in
terms of the composition
%
\coqe!Rows.T (Cols.T (Slice4.T bool))!
%
where \coqe!Slice4.T! denotes a depth of 4 slices. In bitsliced form,
we once again extract out the \coqe!Slice4.T! functor and
\coqe!Double.T! the amount of data processed per run. Fixslicing is
conceptually easier to understand as the diffusion layer is defined as
the combination of an in-register transformation followed by a
90-degree rotation of the matrix of rows and columns. It simply
cancels out after 4 steps.

@@

Of note, the propositions that lead to the bitslicing and fixslicing
correctness theorems hide a somewhat surprising secret: all the proofs
were obtained by \coqe!vm_compute; congruence!. Having defined
\coqe!Rows.T!, \coqe!Cols.T!, \coqe!Slice4.T! and \coqe!Slice8.T! as
records with primitive projections, all the identities boil down to
the specification, bitsliced forms and fixsliced forms to have the
same $\beta$-normal $\eta$-long form through the suitable equivalence
relation!

@@

As part of future work, we intend to extend our formalization to
bridge the gap to machine word. Once again, we expect parametricity to
kick in: we currently have an implementation defined over \coqe!reg32!
(and its supporting operations) which we intend to be equivalent to an
implementation defined over the type of 32 bits machine words (and its
supporting operations).

We also wish to pursue the automated generation of fixsliced code,
eventually implementing a proper simplification engine. In the
process, we ought to develop methods to check for commutativity of the
diffusion layer.

A litmus test to this project will be our ability to deliver a
fixsliced implementation of \AES{}: mark our machine words!

>>>>

We have thus completed our journey toward a rationalized treatment of
bitslicing and fixslicing in their quintessential form, taking the
\Skinny{} cipher as our running example.
%
%% * Other applicative: Gift
%% ** Specification: only thinking: Rows.T Cols.T Slice4.T bool
%% *** rest is rote functionalisation
%% ** Bitslicing: extract Slice4.T out and Double
%% ** Fixslicing: decompose permutation into in-register followed by 90 degree rotation
%% *** cancel out after 4
%% *** -> simpler
%
We have similarly treated the \Gift{} cipher, whose design is in fact at
the origin of the fixslicing
technique~\citep{adominicai:fixslicing-gift}. \Gift{} works over the type
%
\coqe!Rows.T (Cols.T (Slice4.T bool))!
%
where \coqe!Slice4.T! denotes a depth of 4 slices. In bitsliced form,
we once again extract out the \coqe!Slice4.T! functor and
\coqe!Double.T! the amount of data processed per run. Fixslicing is
conceptually easier to understand as the diffusion layer is defined as
the combination of an in-register transformation followed by a
90-degree rotation of the matrix of rows and columns. It simply
cancels out after 4 steps.

@@

Of note, the propositions that lead to the bitslicing and fixslicing
correctness theorems hide a somewhat surprising secret: all the proofs
were obtained by \coqe!vm_compute; congruence!. Having defined
\coqe!Rows.T!, \coqe!Cols.T!, \coqe!Slice4.T! and \coqe!Slice8.T! as
records with primitive projections, all the identities boil down to
the specification, bitsliced forms and fixsliced forms having the
same $\beta$-normal $\eta$-long form through the equivalence
relation!

@@

As part of future work, we intend to extend our formalization to
bridge the gap to machine word. Once again, we expect parametricity to
kick in: we currently have an implementation defined over \coqe!reg32!
(and its supporting operations) which ought to be equivalent to an
implementation defined over the type of 32 bits machine words (and its
supporting operations).
%
We also wish to pursue the automated generation of fixsliced code,
eventually implementing a proper simplification engine. In the
process, we ought to develop methods to check for commutativity of the
diffusion layer.
%
A litmus test to this project will be our ability to deliver a
fixsliced implementation of \AES{}: mark our machine words!
